### 2021-02-20 article/main.tex

\documentclass{article}

\input{header.tex}

% Document starts
\begin{document}

\maketitle

\tableofcontents

\input{intro}
\input{conventions}
\input{partition}
\input{approximation}
\input{codim}
\input{implem}
\input{further}
\appendix
\input{zih}

\bibliography{main}
\bibliographystyle{plain}

\end{document}

%<!-- Local IspellDict: british -->
%<!-- Local IspellPersDict: ~/.ispell-british -->